---
title: "KJ_Pipho_PS_11"
author: "Krista Pipho"
date: "December 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1. 
  Write a function that takes one number and raises it to the power of a second
number

```{r}
# This line defines the function, which takes arguments x and y, then raises x to the power of y
my_powers <- function(x,y) {x^y}
# This line feeds my_powers the aruments 3 and 4, storing the result in 'result'
result <- my_powers(3,4)
# Here we print the result
print(result)
```

Question 2. 
  Write a function that multiples at least three numbers that you pass it as
arguments and that has at least one default value. 

```{r}
# This line defines the function with four arguments, including z which has a default value of 5. The function multiplies all of these arguments together and returns a product. 
my_product <- function(w,x,y,z = 5){w*x*y*z}
# This line calls the function and saves the result as result_2
result_2 <- my_product(2,3,4)
#This line prints result_2
print(result_2)

```

Question 3. 
Call the following with an anonymous function that adds 1 to a variable, x:
evaluate <- function(func, dat){return(func(dat))}

```{r}
# Calls the original function with an anonymous that adds 1 to the argument
evaluate <- function(func, dat){return(func(dat))}
found <- evaluate(function(x){x+1},6)
print(found)

```

Question 4. 
  We accidentally failed to comment our code! Predict what the following code will
do and add commentary to it to let your end user know what it does. (Note: You
will want to confirm that the code works the way you think it should by executing
it in an R chunk but it is more valuable of an exercise if you try to think through it
prior to executing it.)

```{r}
# Generate a vector of length 100 populated with zeros as placeholders
myMeans<-vector("numeric", length=100)
# This should print out 100 zeros if the code worked
print(myMeans)
# Iterate across the vector
for (i in 1:100){
# Generate a random value centered around 0.5. 10 indicates that the program should average 10 values selected with standard probability from around 0.5.
x<-runif(10)
# Overwrite the value you have iterated to with your loop.
myMeans[i]<-mean(x,na.rm=TRUE)
}
#See the results.
print(myMeans)

```

Question 5. 
  Use a higher order function to create a vector that multiplies each element by
itself. Print out the resulting vector to the screen.

```{r}
# Makes a vector with values from 1 to 100
square<-c(1:100)
# Prints that vector
print(square)
# Defines a function that multiplies values by themselves, and feeds it into the higher order function map to call the function on every value in the vector, storing the result in square_vector
square_vector <- Map({function(k) k*k},square)
# Unilists square vector so that it is in regular vector form again
unlist(square_vector)

```

Question 6. 
  Write out three ways of creating a vector that only contains the even numbers
between 0 and 20 (one of these three ways should contain a higher order function,
a second should use a built in function but you should define your own function
for the third way).

```{r}
# This method utilizes the higher order function Filter
small.even.numbers<-Filter(function (x) {x %% 2 == 0}, 1:20)
print(small.even.numbers)

# This method uses a built in function called seq
small_even_2<-seq(2,20, by = 2)
print(small_even_2)

# This is a method I created that provides a third way to generate a list of small even numbers
# Defines func, which takes one number and one vector as arguments 
func <- function(x,a){
  # Makes a loop with as many itterations as the numerical argument
  for (i in 1:x)
  # Only appends the number of the itteration to the vector if that number is even by mod 2
  {if (i %% 2 == 0) {a <- c(a,i)}}
# Returns the now full vector
return(a)}
# Initializes an empty vector
a <- c()
# Calls func with the empty vector and 20
small_3 <- func(20,a) 
# Prints the result
print(small_3)
```
